Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-06-09 paper.Rmd
add model 2 model comparison plots in appendix
also drop in last rep_nosmooth run results

```{r benefit, fig.height=3.5, fig.width=5.5, out.width = ".8\\linewidth", fig.align = "center", fig.cap = "Gain from using smoothed ROC to calculate AUC, for each model reported in Tables 3 and 4 (B\\&S Tables 1 and 2)"}
table1_benefit <- read_csv("data/table1-smooth-benefit.csv") 
table2_benefit <- read_csv("data/table2-smooth-benefit.csv")

@@

>>>>

```{r benefit-plot, fig.height=3.5, fig.width=5.5, out.width = ".8\\linewidth", fig.align = "center", fig.cap = "Gain from using smoothed ROC to calculate AUC, for each model reported in Tables 3 and 4 (B\\&S Tables 1 and 2)"}
table1_benefit <- read_csv("data/table1-smooth-benefit.csv") 
table2_benefit <- read_csv("data/table2-smooth-benefit.csv")

@@

## Model to model comparison plots

Figures \ref{fig:table1-plot} and \ref{fig:table2-plot} replicate the information in Tables \ref{tab:table1-nosmooth} and \ref{tab:table2-fixed} in a way makes the comparison of the escalation model to the alternative models easier. Each facet shows the escalation AUC-ROC for all model settings (the rows in Table \ref{fig:table1-plot} and base specification for all models in Table \ref{tab:table2-fixed}) on the left, and the AUC-ROC for an alternative model on the right, with a connecting line. If the lines slope up to the right, the alternative model is better.

```{r table1-plot, fig.height=5, fig.width=8, out.width = ".9\\linewidth", fig.align = "center", fig.cap = "Escalation to alternative comparisons for the ICEWS models (B\\&S Table 1)"}
table1_nosmooth <- read_csv("data/table1-nosmooth.csv")

icews_lvls <- c("Escalation", "Quad", "Goldstein", "CAMEO", "Average")

df <- table1_nosmooth %>%
  rename(Spec = Model) %>%
  pivot_longer(Escalation:Average, names_to = "Model", values_to = "auc_roc") %>%
  mutate(Model = factor(Model, levels = icews_lvls)) %>%
  nest(data = c(Spec, auc_roc))

pairs <- full_join(
  df %>% filter(Model!="Escalation"),
  df %>% filter(Model=="Escalation") %>% rename(data2 = data) %>% select(horizon, data2),
  by = c("horizon")
) %>%
  pivot_longer(data:data2, names_to = "xname", values_to = "data") %>%
  mutate(xname = ifelse(xname=="data", as.character(Model), "Escalation"),
         xname = factor(xname, levels = icews_lvls))

pairs %>%
  unnest(data) %>%
  ggplot(aes(x = xname, y = auc_roc)) +
  facet_grid(horizon ~ Model, scales = "free_x") +
  geom_point() +
  geom_line(aes(group = Spec)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "", y = "AUC-ROC (empirical)")
```

```{r table2-plot, fig.height=5, fig.width=8, out.width = ".9\\linewidth", fig.align = "center", fig.cap = "Escalation to alternative comparisons for the structural models (B\\&S Table 2)"}
table2_nosmooth <- read_csv("data/table2-nosmooth.csv")

model_lvls <- colnames(table2_nosmooth)[-c(1:2)]

df <- table2_nosmooth %>%
  rename(Spec = Model) %>%
  pivot_longer(`Escalation Only`:`PITF Only`, names_to = "Model", values_to = "auc_roc") %>%
  mutate(Model = factor(Model, levels = model_lvls)) %>%
  nest(data = c(Spec, auc_roc))

pairs <- full_join(
  df %>% filter(Model!="Escalation Only"),
  df %>% filter(Model=="Escalation Only") %>% rename(data2 = data) %>% select(horizon, data2),
  by = c("horizon")
) %>%
  pivot_longer(data:data2, names_to = "xname", values_to = "data") %>%
  mutate(xname = ifelse(xname=="data", as.character(Model), "Escalation Only"),
         xname = factor(xname, levels = model_lvls))

pairs %>%
  unnest(data) %>%
  ggplot(aes(x = xname, y = auc_roc)) +
  facet_grid(horizon ~ Model, scales = "free_x") +
  geom_point() +
  geom_line(aes(group = Spec)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(x = "", y = "AUC-ROC (empirical)")
```
